---
title: "Getting started"
author: 
  - name: Regina H. Reynolds
    affiliation:
      - UCL
    email: regina.reynolds.16@ucl.ac.uk
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('rprojecttemplate')`"
vignette: >
  %\VignetteIndexEntry{Introduction to rprojecttemplate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# Installation

To access the template on RStudio first install `rprojecttemplate` with:

``` {r, eval = F}

devtools::install_github("RHReynolds/rprojecttemplate")

```

# Quick start to using to `rprojecttemplate`

```{r "start", message=FALSE}
```

This [R] package was made for personal use, and aims to create an R project directory primed for data analysis. It has a simple focus on:

1. Creating a standardized project folder structure with a few template files
needed for beginning a data analysis project
2. Use of practices that make it easier for the project to be reproducible and open

The structure makes use of the existing and established packages ([`devtools`](https://devtools.r-lib.org/), [`usethis`](https://usethis.r-lib.org/), and [`renv`](https://rstudio.github.io/renv/articles/renv.html)) and is intended for use with RStudio.

# Usage


# Acknowledgements
This package was inspired by the [`prodigenr`](https://github.com/rostools/prodigenr) package.

# Session info

<details>
  <summary>Show/hide</summary>
  
```{r session-info, echo = FALSE}
library("sessioninfo")
options(width = 120)
session_info()
```

</details> 
