---
title: "Getting started"
author: 
  - name: Regina H. Reynolds
    affiliation:
      - UCL
    email: regina.reynolds.16@ucl.ac.uk
output: 
  BiocStyle::html_document:
    self_contained: yes
    toc: true
    toc_float: true
    toc_depth: 2
    code_folding: show
date: "`r doc_date()`"
package: "`r pkg_ver('rprojecttemplate')`"
vignette: >
  %\VignetteIndexEntry{Introduction to rprojecttemplate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    crop = NULL
)
```

# Installation

To access the template on RStudio first install `rprojecttemplate` with:

``` {r, eval = F}
devtools::install_github("RHReynolds/rprojecttemplate")
```

# Quick start to using to `rprojecttemplate`

This `R` package was made for personal use, and aims to create an R project directory primed for data analysis that can be hosted either on GitHub or GitLab. It has a simple focus on:

1. Creating a standardised project folder structure with a few template files
needed for beginning a data analysis project
2. Use of practices that make it easier for the project to be reproducible and open

The structure makes use of the existing and established packages ([`devtools`](https://devtools.r-lib.org/), [`usethis`](https://usethis.r-lib.org/), and [`renv`](https://rstudio.github.io/renv/articles/renv.html)) and is intended for use with RStudio.

# Usage

## My typical workflow

Personally, I believe all data analyses should be hosted on Github/GitLab -- both for reproducibility, but also for your own version control. 

Pre-publication, these can be hosted as private repositories and made public on release of a paper. Just remember to think about whether your repository contains any sensitive information (e.g. patient-identifiable information or IP addresses for private servers).

My typical workflow is:

1. Create a GitHub/GitLab repository on the GitHub/GitLab website, as detailed in this [GitHub guide](https://docs.github.com/en/get-started/quickstart/create-a-repo) or [GitLab guide](https://docs.gitlab.com/ee/user/project/working_with_projects.html#create-a-project).
2. Start a new project in RStudio via "File -> New Project", then "Version Control --> Git".
3. Set up my analysis project with `rprojecttemplate::setup_project()`.
4. Make an initial git commit with analysis template.
5. Set up [`renv`](https://rstudio.github.io/renv/articles/renv.html) for reproducibility.

## Getting started

### GitHub-hosted project

There are two ways of creating a new project: using the R console or using the RStudio "New Project" menu option.

Through the console, use the `setup_project()` command. For example:

```{r manuscript-proj}
library(rprojecttemplate)
# Create a temporary folder using the fs package
new_project_path <- fs::path_temp("example")
rprojecttemplate::setup_project(new_project_path, type = "github")
```

Please note that if no `type` is provided to `setup_project()` the assumption is that the project will be hosted on GitHub, and a template created accordingly.

This then creates a directory tree, with template files for starting your analysis. Open the newly created project via the `.Rproj` file. 

For the RStudio approach, go to "File -> New Project", then "New directory" and find the `rprojecttemplate` project in the list (it should be listed as "Data analysis project using rprojecttemplate").

The resulting file structure should look something like this:

```{r folder-tree, echo=FALSE, comment="", results="markup"}
rprojecttemplate:::viz_project_tree(new_project_path)
```

`README` files are contained within each project and in each folder that explains a bit more about what each folder and file are used for, which is briefly described here:

- `DESCRIPTION`: Is a standard file that includes metadata about your project, in a machine readable format, and that also stores a list of the R packages your project depends on. An example function is included. Consider using [roxygen-style](https://roxygen2.r-lib.org/) 
documentation of your functions.
- `.Rproj`: Is a standard file used by RStudio to set some R Project specific settings.
- `R/`: Should contain the functions called in `docs/` and `scripts/`.
- `docs/`: Should contain all `.Rmd`s and their corresponding `.html`s describing analyses performed for this project. An example `.rmd` is included along with a table of contents (`index.md`), which describes the purpose of each analysis and can be used as a landing page if you choose to host your `docs/` on [GitHub pages](https://pages.github.com/).
- `raw_data/`: Folder should either:
    - Contain the raw, unprocessed data used specifically for each analysis. 
    - In those situations where raw data files are too large to be loaded to GitHub, consider:
        - Uploading those files to a repository like zenodo or figshare (assuming they are do not contain patient-sensitive information).
        - If the files are publicly available, include either a script with commands to download these files can be a text file describing where and how to download these files.
- `results/`: Results generated from any analyses/scripts that are part of the project should be stored here as data files.
- `scripts/`: Contains analysis scripts. Each script should contain a one-line description and should be named such that it corresponds to an `.Rmd` in `docs/`. Consider using numbering and lettering to denote the sequence in which scripts should be run.

The template I use for my analysis `.rmd`s can be installed using my [`rmdplate` package](https://github.com/RHReynolds/rmdplate). Alternatively, consider making your own template.

### GitLab-hosted project

As with a GitHub-hosted project, there are two ways of creating a new project: using the R console or using the RStudio "New Project" menu option.

Through the console, use the `setup_project()` command, but specify `type = "gitlab"`. For example:

```{r manuscript-proj-gitlab}
# Create a temporary folder using the fs package
new_project_path <- fs::path_temp("example-gitlab")
rprojecttemplate::setup_project(new_project_path, type = "gitlab")
```

This then creates a directory tree, with template files for starting your analysis. Open the newly created project via the `.Rproj` file. 

For the RStudio approach, go to "File -> New Project", then "New directory" and find the `rprojecttemplate` project in the list (it should be listed as "Data analysis project for gitlab using rprojecttemplate").

The resulting file structure should look something like this:

```{r folder-tree-gitlab, echo=FALSE, comment="", results="markup"}
rprojecttemplate:::viz_project_tree(new_project_path)
```

The main differences between this GitLab-friendly template and the GitHub-friendly template are:

- `.gitlab-ci.yml` - basic template for CI that is required to add a Pages site
- `index.rmd` - the landing page for GitLab pages needs to within the main repository (as opposed to `docs/`). Remember to knit this page everytime an edit is made, using `devtools::build_rmd("index.rmd")`.

# Acknowledgements
This package was heavily inspired by the [`prodigenr`](https://github.com/rostools/prodigenr) package, thus be sure to cite it if you use this `R` package.

# Session info

<details>
  <summary>Show/hide</summary>
  
```{r session-info, echo = FALSE}
library("sessioninfo")
options(width = 120)
session_info()
```

</details> 
